---
title: "Statistical Inference Project"
author: "Julie Foster"
date: "October 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Part 1: Simulation Exercise
 The first part of this project investigates the exponential 
 distribution in 
 R and compares it with the Central Limit Theorem. The exponential 
 distribution is simulated in R with rexp(n, lambda) where 
 lambda is the rate parameter. The mean of exponential distribution 
 is 1/lambda and the standard deviation is also 1/lambda. Set 
 lambda = 0.2 for all of the simulations. 
```{r}
lambda = 0.2
mn = 1/lambda
sd = 1/lambda
```

Here is the simulation for 1000 averages of 40 exponentials and set
to the variable "exp_mns".

```{r}
n = 1000
exp_mns = NULL
for (i in 1:1000) exp_mns = c(exp_mns, mean(rexp(40, lambda)))
```

### 1. Show the sample mean and compare it to the theoretical mean of the  distribution.

```{r}
sample_mean <- mean(exp_mns)
sample_mean

pop_mean <- mn
pop_mean
```

"sample_mean" is the mean of sample values collected.
"population_mean"" is the mean of all the values in the population.
As is shown here, the sample mean is a good estimate of the population mean.

### 2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

```{r}
sample_var <- sd(exp_mns)/sqrt(n)
sample_var

pop_var <- sd^2/1000
pop_var
```

Similarly to how the sample mean compares with the population mean, the sample variance is also an unbiased estimator of the population variance.

### 3. Show that the distribution is approximately normal.

```{r}
data <- rexp(n, lambda)
par(mfrow=c(1,2))
hist(data, probability=TRUE, main="Random Exponentials", 
     xlab="Exponentials", col="steelblue", las=1, breaks=30)
lines(density(data), lwd=2)
abline(v=pop_mean, lwd=3, col="red")
hist(exp_mns, probability=TRUE, main="Averages of Random
     Exponentials", xlab="Averages of Exponentials", col="coral",
     las=1, breaks=30)
lines(density(exp_mns), lwd=2)
abline(v=sample_mean, lwd=3, col="red")
```

Here are two plots comparing the distribution of a thousand random 
exponentials with the distribution of the averages of a thousand 
random exponentials. The second plot shows the distribution of the
averages, which illustrates that the distribution is approximately 
normal, with the sample mean (shown as the red vertical line) as the 
place where the distribution is clustering.



# Part 2: Inferential Data Analysis 

The second part of this project is the analysis of the ToothGrowth 
dataset.  The sample size of this dataset is relatively small and 
therefore t.tests will be performed to analyze the confidence
intervals.

First the dataset is loaded and some basic exploration of the data
is performed:

```{r}
data(ToothGrowth)
dim(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
plot(ToothGrowth)
```

The data shows two supplement groups: OJ and VC, that do not look to
be paired up.  Variance is the same in both groups because the
sample size is the same, each with a size of 30.
There are also three dosage groups: 0.5, 1.0, and 2.0.

Looking at the results from the t.test function to show tooth length
vs. supplement:

```{r}
t.test(len ~ supp, paired = FALSE, var.equal = TRUE, data = ToothGrowth)
```

Comparing tooth length by dose, looking at the different pairs of dose values.

```{r}
t.test(len~dose, data=ToothGrowth[ToothGrowth$dose==0.5|ToothGrowth$dose==1.0,])
t.test(len~dose, data=ToothGrowth[ToothGrowth$dose==1.0|ToothGrowth$dose==2.0,])
t.test(len~dose, data=ToothGrowth[ToothGrowth$dose==0.5|ToothGrowth$dose==2.0,])

```
## Conclusions
Assumptions:  

* the sample is representative of the total population  
* the distribution of the sample means follows the Central Limit 
Theorem  

Results:    

* For the tooth length vs supplement type:  
    The confidence interval includes the value zero and the p-value is
    shown as 0.06 which is greater than 0.05, so therefore the 
    supplement types seem to have no impact on tooth length.   

* For the tooth length vs dosage level:
    The p-value of each of these tests are all essentially zero and
    the confidence intervals of these tests do not cross zero, so 
    therefore the dosage levels to appear to have an impact on 
    tooth length.

